---
title: "Twist Capture Methyl-Seq QC Report"
date: "`r Sys.Date()`"
encoding: UTF-8 
params:
    panel: NULL
    coverage: NULL
    cell_counts: NULL
    meth: NULL
    score_sites: NULL
format:
  html:
    fig-cap-location: bottom
    df-print: paged
    fig-width: 10
    fig-height: 8
    embed-resources: true
    toc: true
    cache: false
    warning: false
---

```{r, include=FALSE}
#| label: load-data

library(data.table)
library(ggplot2)
library(gridExtra)
library(reshape2)
library(RColorBrewer)
library(knitr)
library(impute)
library(meffonym)

panel = fread(params$panel)
coverage = fread(params$coverage)
cell_counts = fread(params$cell_counts)
meth = fread(params$meth)
score_sites = fread(params$score_sites)

stopifnot(identical(colnames(meth), colnames(coverage)))
stopifnot(all(c("chr","start","end") %in% colnames(meth)))
stopifnot(ncol(panel) == 3)

colnames(panel) = c("chr","start","end")
colnames(cell_counts)[1] = "cell_type"
colnames(score_sites)[1] = "score"

setkey(panel, chr, start, end)
setkey(meth, chr, start, end)
setkey(coverage, chr, start, end)
panel_meth = foverlaps(meth, panel, type="within", nomatch=NULL)
panel_coverage = foverlaps(coverage, panel, type="within", nomatch=NULL)

panel = data.frame(panel)
coverage = data.frame(coverage)
cell_counts = data.frame(cell_counts)
meth = data.frame(meth)
score_sites = data.frame(score_sites)
panel_meth = data.frame(panel_meth)
panel_coverage = data.frame(panel_coverage)

meth = meth[,setdiff(colnames(meth),c("chr","start","end"))]
panel_meth = panel_meth[,colnames(meth)]
coverage = coverage[,colnames(meth)]
panel_coverage = panel_coverage[,colnames(meth)]

stopifnot(nrow(meth)==nrow(coverage))
stopifnot(nrow(panel_meth)==nrow(panel_coverage))
stopifnot(all(colnames(meth) %in% colnames(cell_counts)))

rownames(cell_counts) = cell_counts$cell_type
cell_counts$cell_type = NULL

samples = data.frame(
    id=colnames(meth),
    name=sub("_[^_]*", "", colnames(meth)))
rownames(samples) = samples$id
```

## Overview

A CpG site is considered **measured** or **covered** in this dataset
if it is contained in at least `mincov` (10) reads for at least
50% of the samples. 

A CpG site is called **targeted** if probes were designed to
to capture DNA fragments containing that site.

```{r}
#| echo: false
#| label: panel-overview

overview = rbind(
    data.frame(
        metric="covered CpG sites", 
        value=nrow(meth)),
    data.frame(
        metric="targeted genomic regions",
        value=nrow(panel)),
    data.frame(
        metric="targeted and covered CpG sites",
        value=nrow(panel_meth)),
    data.frame(
        metric="number of samples analysed",
        value=ncol(panel_meth)),
    data.frame(
        metric="coverage threshold (mincov)",
        value=10))
kable(overview)
```
## Methylation distributions

DNA methylation distributions are shown by sample.

```{r}
#| echo: false
#| label: methylation-density
#| code-fold: false
#| fig-cap: "**Figure 1**: Methylation distributions by sample"

long = melt(panel_meth, variable.name="sample", value.name="methylation")
long$sample = samples[long$sample, "name"]
long = na.omit(long)

ggplot(long, aes(x=methylation, fill=sample, colour=sample)) + 
    geom_density(alpha = 0.25, linewidth = 0.6) +
    labs(x = "methylation", 
         y = "density") + 
    theme(legend.position = "right")
```

## Cell count estimates

Cell counts are deconvoluted DNA methylation levels and a blood cell
type reference
([Loyfer, et al. Nature, 2023](https://pubmed.ncbi.nlm.nih.gov/36599988/)).


```{r}  
#| echo: false
#| code-fold: false
#| label: cell-counts
#| fig-cap: "**Figure 2**: Estimated cell count distributions across samples"

long = melt(
    setDT(data.frame(t(cell_counts))),
    variable.name="cell_type",
    value.name="cell_count")

med = aggregate(cell_count ~ cell_type, data=long, median, na.rm=T)
med$cell_count = round(med$cell_count,2)

ggplot(long, aes(x=cell_type, y=cell_count, fill=cell_type)) +
    geom_violin(
        alpha = 0.25, 
        scale = "width",
        trim = FALSE,
        position = position_dodge(width = 0.5)) + 
    theme_minimal(base_size = 15) +  
    labs(x="cell type", y="cell counts") +
    geom_text(
        data=med,
        aes(x=cell_type, y=cell_count, label=paste0("med=",cell_count)),
        size=3)+
    theme(
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.title = element_text(size = 10),
        plot.title = element_text(size = 12, hjust = 0.5),
        legend.position = "none")
```

## Distributions of read depths per sample

Distribution of read counts per sample covering targeted CpG sites. 

```{r}
#| echo: false
#| label: depth-distribution
#| code-fold: false
#| fig-cap: "**Figure 3**: Read depth per sample"

long = melt(setDT(panel_coverage), variable.name="sample", value.name="reads")
long$sample = samples[long$sample, "name"]
long$reads[is.na(long$reads)] = 0

med_reads = aggregate(reads ~ sample, data=long, median)
med_reads = med_reads[order(med_reads$reads, decreasing=T),]
med_reads$reads = round(med_reads$reads,2)

long$sample = factor(
    as.character(long$sample),
    levels=med_reads$sample,
    ordered=T)

ggplot(long, aes(x=sample, y=reads, fill=sample)) +
    geom_violin(trim=F, alpha=0.7, position=position_dodge(width=0.8)) +
    geom_text(
        data=med_reads, 
        aes(x=sample, y=reads, label=paste0("med=",reads)),
        size=3)+
    theme_minimal(base_size = 15) +
    labs(x="sample", y="read depth") + 
    theme(
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        axis.title = element_text(size = 10),
        plot.title = element_text(size = 12, hjust = 0.5),
        legend.position = "none")
```

## Pairwise correlations between samples

```{r}
#| echo: false
#| label: methylation-correlation
#| error: true
#| fig-cap: "**Figure 4**: Pairwise correlations between samples"

r = cor(panel_meth, use="p")
diag(r) = NA
r = na.omit(as.vector(r))

ggplot(data.frame(r=r), aes(x=r)) +
    geom_density(fill="blue", alpha=0.5) +
    labs(x="Pearson's correlation", y="density") +
    theme_minimal()
```

## Principal components of methylation vs read depth

```{r}
#| echo: false
#| label: pca
#| error: true
#| fig-cap: "**Figure 5**: Principal components of methylation vs read depth "

set.seed(20250715)
n_features = min(100000, nrow(panel_meth))
subset_meth = panel_meth[sample(nrow(panel_meth), n_features),]
subset_meth = impute.knn(t(subset_meth))$data
subset_meth = subset_meth[,apply(subset_meth,2,var,na.rm=T) > 1e-16]
ret = prcomp(subset_meth, scale=T, center=T)
pcs = ret$x
rownames(pcs) = samples[rownames(pcs), "name"]
pcs = data.frame(sample=rownames(pcs), pcs)

med_coverage = apply(panel_coverage, 2, median, na.rm=T)
names(med_coverage) = samples[names(med_coverage), "name"]
pcs$reads = med_coverage[rownames(pcs)]

pairs = t(combn(paste0("PC",1:3),2))
for (i in 1:nrow(pairs)) {
    p = ggplot(pcs, aes_string(x=pairs[i,1], y=pairs[i,2], color="reads")) +
        geom_point() + 
        geom_text(aes(label=sample,vjust=-0.5,hjust=0.5),size=3) +
        labs(color="median\nread\ndepth") + 
        theme_minimal() +
        scale_color_gradient(low="blue", high="red")
    print(p)
}

ret$pct = ret$sdev^2/sum(ret$sdev^2)*100
npcs = min(10,length(ret$pct))
ggplot(
    data.frame(pc=colnames(ret$x)[1:npcs], pct=ret$pct[1:npcs]),
    aes(x=pc, y=pct)) +
    geom_bar(stat="identity", fill="steelblue") +
    labs(title="Scree plot", x="principal components", y="% variance") +
    theme_minimal()
```

## Covered CpG sites in published methylation score

Published DNA methylation scores
for a wide range of phenotypes and exposures
have been implemented in the
[meffonym R package](https://github.com/perishky/meffonym).
Below we provide some information about
the CpG sites in these scores that are covered in this dataset.

```{r}
#| echo: false
#| label: scores
#| error: true
#| fig-cap: "**Figure 6**: Percent of measured sites in each methylation score"

ggplot(
    score_sites,
    aes(pct)) +
    geom_density(alpha=0.25, linewidth=0.6) + 
    labs(x = "% sites in methylation scores", 
         y = "density") + 
    theme(legend.position = "right")
```

Below are the `r sum(score_sites$pct > 80)`
of `r length(meffonym.models())`
available scores for which at least 80% of sites are covered. 

```{r}
#| echo: false
#| label: scores_over_80
#| error: true
kable(score_sites[score_sites$pct > 80,])
```

