---
title: "Twist Capture Methyl-Seq QC Report"
author: "."
date: "`r Sys.Date()`"
encoding: UTF-8 
params:
    panel: NULL
    coverage_matrix: NULL
    cell_counts: NULL
    methylation_matrix: NULL
format:
  html:
    fig-cap-location: bottom
    df-print: paged
    fig-width: 10
    fig-height: 8
    embed-resources: true
    toc: true
    cache: false
    warning: false
---

```{r, include=FALSE}
#| label: load-data

library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(reshape2)
library(RColorBrewer)
library(glue)

panel = fread(params$panel)
coverage_matrix = fread(params$coverage_matrix)
cell_counts = fread(params$cell_counts)
methylation_matrix = fread(params$methylation_matrix)

colnames(cell_counts)[1] <- "cell_type"
setkey(panel, chr, start, end)
setkey(methylation_matrix ,  chr, start, end)
input_data         <- foverlaps(methylation_matrix, panel, type = "within", nomatch = 0L)
input_coverage     <- foverlaps(coverage_matrix, panel, type = "within", nomatch = 0L)
input_data         <- input_data %>% select(-matches("chr|start|end|details|source"))
methylation_matrix <- methylation_matrix %>% select(-matches("chr|start|end"))
coverage_matrix    <- coverage_matrix %>% select(-matches("chr|start|end"))
input_coverage     <- input_coverage  %>% select(-matches("chr|start|end|details|source|i.start|i.end"))
```

* <em>All metrics and plots in this report refer to the predefined CpG panel unless explicitly noted otherwise.</em>

## <sup>Overview Table</sup>

```{r}
#| echo: false
#| label: panel-overview
#| results: asis

# Numbers

total_cpg_all       <- nrow(methylation_matrix)
total_cpg_panelfile <- nrow(panel)
cpg_after_filter    <- nrow(input_data)
num_samples         <- ncol(input_data)
coverage_threshold  <- 10

#Descriptions

metrics <- c(
  "Covered CpG sites",
  "Targeted CpG sites",
  "Targeted and covered CpG sites",
  "Number of samples analysed",
  "Coverage threshold (mincov)"
)

descriptions <- c(
  "The number of CpG sites in the human genome for which at least 50% of the samples each have at least 'mincov' reads covering the site.",
  "The number of CpG sites inside panel regions",
  "The number of covered CpG sites inside panel regions",
  "",
  ""
)

values <- format(
  c(total_cpg_all,
    total_cpg_panelfile,
    cpg_after_filter,
    num_samples,
    coverage_threshold),
  big.mark = ","
)

row_html <- vapply(seq_along(metrics), function(i) {
  glue('  <tr>
    <td style="padding:6px 10px;">
      {metrics[i]}<br>
      <span class="subdesc">{descriptions[i]}</span>
    </td>
    <td style="padding:6px 10px; text-align:right;">{values[i]}</td>
  </tr>')
}, character(1))

# CSS codes

css <- '
<style>
  /* Table size & typography */
  table.metric-table {
    width:90%;
    border-collapse:collapse;
    font-size:0.95rem;
    margin:1rem auto;
  }

  /* Subtle gray borders */
  .metric-table,
  .metric-table th,
  .metric-table td {
    border:1px solid #d8d8d8;
  }

  /* Header row – light grey background, darker text */
  .metric-table th {
    background:#f2f2f2;
    color:#000000;
    text-align:left;
    padding:6px 10px;
    font-weight:600;
  }

  /* Data cells – white + gentle zebra striping */
  .metric-table td {
    background:#ffffff;
    color:#333333;
    padding:6px 10px;
  }
  .metric-table tr:nth-child(even) td { background:#fafafa; }

  /* Description under each metric – darker & semi-bold */
  .metric-table .subdesc {
    font-style:italic;
    font-size:0.85em;
    color:#4a4a4a;
    font-weight:600;
  }
</style>
'

#HTML out

html_out <- glue('{css}
<table class="metric-table">
  <thead>
    <tr><th>Metric</th><th style="text-align:right;">Value</th></tr>
  </thead>
  <tbody>
{paste(row_html, collapse = "\n")}
  </tbody>
</table>
')

cat(html_out)
```
## <sup>Methylation Distribution Plot</sup>

* This plot shows the distribution of DNA methylation levels across multiple samples. 
* The x-axis represents methylation values from 0 (unmethylated) to 1 (fully methylated), while the y-axis shows the density of CpG sites. 
* Each line corresponds to a sample, with color shading indicating overlapping densities. 
* The bimodal pattern, with peaks near 0 and 1, is typical for methylation data, highlighting unmethylated and highly methylated CpG regions. 
* This plot captures methylation variability across samples.

```{r}
#| echo: false
#| label: methylation-density
#| code-fold: false
#| fig-cap: "**Figure 1**: Density Plot of Methylation for Samples."

methylation_data_long <- input_data %>%
  pivot_longer(cols = everything(), names_to = "Sample", values_to = "Methylation")

ggplot(methylation_data_long,
       aes(x = Methylation, fill = sub(".*_(S\\d+)$", "\\1", Sample),
           colour = sub(".*_(S\\d+)$", "\\1", Sample))) +
  geom_density(alpha = 0.25, linewidth = 0.6) +
  labs(title = "Density Plot of Methylation for Samples",
       x = "Methylation", 
       y = "Density",
       fill = "Sample", 
       colour = "Sample") + 
  theme(legend.position = "right") 
```
## <sup>Estimate cell counts</sup>

* The plot shows data density with labels for Q1, median, and Q3.
* These cell-type proportions are inferred from genome-wide methylation patterns and do not rely on the targeted CpG panel. 
* <em>In other words, the estimates reflect the overall cellular composition of each sample, independent of panel capture efficiency or on-target depth.</em>

```{r}  
#| echo: false
#| code-fold: false
#| label: cell-counts
#| fig-cap: "**Figure 2**: The violin plot displays the distribution of estimated cell counts across cell types."

getPalette <- colorRampPalette(brewer.pal(8, "Set2"))
data <- cell_counts %>%
  pivot_longer(cols = -cell_type, names_to = "Sample", values_to = "Value") %>%
  rename(CellType = cell_type)

iqr_data <- data %>%
  group_by(CellType) %>%
  summarize(
    Q1 = quantile(Value, 0.25),
    Median = median(Value),
    Q3 = quantile(Value, 0.75))
violin_plot <- ggplot(data, aes(x = CellType, y = Value, fill = CellType)) +
  geom_violin(alpha = 0.25, 
   scale    = "width",
   trim     = FALSE,
   position = position_dodge(width = 0.5)) + 
  scale_fill_manual(values = getPalette(length(unique(data$CellType)))) +  
  geom_text(data = iqr_data, aes(x = CellType, y = Q3, label = paste0("Q3: ", round(Q3, 2))), 
            position = position_nudge(x = 0.2), size = 3) + 
  geom_text(data = iqr_data, aes(x = CellType, y = Median, label = paste0("Median: ", round(Median, 2))), 
            position = position_nudge(x = 0.2), size = 3) +  
  geom_text(data = iqr_data, aes(x = CellType, y = Q1, label = paste0("Q1: ", round(Q1, 2))), 
            position = position_nudge(x = 0.2), size = 3) +  
  theme_minimal(base_size = 15) +  
  labs(
    title = "Violin Plot: Distribution of Cell Types",
    x = "Cell Type",
    y = "Value"
  ) +
  theme(
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, hjust = 0.5),
    legend.position = "none"  
  )
print(violin_plot)
```

## <sup>Depth Distribution</sup>

* This plot illustrates the distribution of depth for each sample, with Samples represented along the x-axis and depth values on the y-axis.
* Side-by-side view: Two boxes per sample—Panel and Total—compare on-target depth to genome-wide depth at a glance.
* Sorted by quality: Samples are arranged by median Panel depth, putting the best on the left and weak outliers on the right.
* Quick red-flags: Panel boxes should be higher and tighter; lower or wider ones signal poor capture or uneven depth.

```{r}
#| echo: false
#| label: depth-distribution
#| code-fold: false
#| fig-cap: "**Figure 3**:  Comparison of Depth per Sample Between Panel CpG sites and Total CpG sites."

depth_long <- bind_rows(
  input_coverage   %>% 
    mutate(Set = "Panel") %>%          
    pivot_longer(cols = -Set,
                 names_to = "Sample",
                 values_to = "Depth"),
  coverage_matrix %>% 
    mutate(Set = "Total") %>% 
    pivot_longer(cols = -Set,
                 names_to = "Sample",
                 values_to = "Depth")
)

depth_long <- depth_long %>%
  mutate(SampleLabel = sub(".*_(S\\d+)", "\\1", Sample))

sample_order <- depth_long %>% 
  group_by(SampleLabel) %>% 
  summarise(mean_cov = mean(Depth, na.rm = TRUE), .groups = "drop") %>% 
  arrange(desc(mean_cov)) %>% 
  pull(SampleLabel)

depth_long$SampleLabel <- factor(depth_long$SampleLabel,
                                    levels = sample_order)

ymax <- max(depth_long$Depth, na.rm = TRUE)

ggplot(
  depth_long,
  aes(x = SampleLabel, y = Depth, fill = Set)
) +
  geom_violin(
    trim     = FALSE,
    scale    = "width",
    width    = .8,
    alpha    = .70,
    position = position_dodge(width = .8)
  ) +
  stat_summary(
    fun      = median,
    geom     = "point",
    size     = .5,
    colour   = "black",
    position = position_dodge(width = .8)
  ) +
  scale_y_continuous(
    limits = c(0, ymax),
    expand = expansion(mult = c(0, .05))
  ) +
  scale_fill_manual(values = c(Panel = "#2c7bb6", Total = "#d7191c")) +
  labs(
    title = "Depth per Sample: Panel vs Total",
    x     = "Sample",
    y     = "Depth",
    fill  = "Set"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x        = element_text(angle = 90, vjust = 0.5, hjust = 1),
    panel.grid.major.x = element_blank()
  ) 
```

## <sup>Sample-wise Correlation</sup>
```{r}
#| echo: false
#| label: methylation-correlation
#| error: true
#| code-fold: false

generate_correlation_plots <- function(data) {  
  # Calculate correlation matrices
  cor_matrix_no_standardize <- cor(data, use = "pairwise.complete.obs")
  meth_data_scaled_by_cpg <- t(scale(t(data)))  # Standardize CpG sites row-wise
  cor_matrix_standardized <- cor(meth_data_scaled_by_cpg, use = "pairwise.complete.obs")
  
  # Melt correlation matrices for plotting
  cor_melt_no_standardize <- melt(cor_matrix_no_standardize)
  cor_melt_standardized <- melt(cor_matrix_standardized)
  
  # Plot density of correlations without standardization
  p1 <- ggplot(cor_melt_no_standardize, aes(x = value)) +
    geom_density(fill = "blue", alpha = 0.5) +
    labs(title = paste("Density Plot Without Standardizing CpG Sites"), 
         x = "Correlation", y = "Density") +
    theme_minimal()
  
  # Plot density of correlations with standardization
  p2 <- ggplot(cor_melt_standardized, aes(x = value)) +
    geom_density(fill = "red", alpha = 0.5) +
    labs(title = paste("Density Plot With Standardizing CpG Sites"), 
         x = "Correlation", y = "Density") +
    theme_minimal()
  
  # Print the plots
   return(list(p1 = p1, p2 = p2))
} 
  # Generate correlation plots for full dataset
 plots <- generate_correlation_plots(input_data)
```
* This plot shows the distribution of pairwise correlations between samples without standardizing the CpG sites. 
* The plot provides an overview of the raw correlation data.

```{r}
#| echo: false
#| label: raw-methylation-correlations
#| error: true
#| fig-cap: "**Figure 4**: Density Plot Without Standardizing CpG Sites."

# Print the density plot for correlations without standardization
print(plots$p1)
```
* This plot shows the distribution of pairwise correlations between samples after standardizing the CpG sites. 
* Standardization normalizes CpG sites across samples, offering a clearer view of relative variations in the data.

```{r}
#| echo: false
#| label: standardized-methylation-correlations
#| error: true
#| fig-cap: "**Figure 5**: Density Plot With Standardizing CpG Sites."
print(plots$p2)
```

## <sup>Coverage density per sample</sup>
 
```{r}
#| echo: false
#| label: coverage-density
#| error: true
#| fig-cap: "**Figure 6**: Coverage density per sample"

n_cpg_sites <- nrow(coverage_matrix)
n_samples   <- ncol(coverage_matrix)
coverage_data_long <- coverage_matrix %>%
  pivot_longer(
    cols = everything(),
    names_to = "Sample",
    values_to = "Coverage"
  )
cat("Number of CpG sites:", n_cpg_sites, "\n")
cat("Number of samples:", n_samples, "\n")

cov <- ggplot(coverage_data_long, aes(x = Coverage, fill = Sample)) +
  geom_density(alpha = 1) +
  labs(
    title = "Density Plot of Coverage for Samples",
    x = "Coverage",
    y = "Density"
  ) +
  theme_minimal() +
  xlim(0, 600) +
  theme(legend.position = "none")
```
