---
title: "DNA Methylation: Bisulfite Sequencing Panel Report"
author: "."
date: "`r Sys.Date()`"
params:
  dataset: "."      
  panel_data: NULL      
  file_list: "qc_files.csv"
format:
  html:
    fig-cap-location: bottom
    df-print: paged
    fig-width: 10
    fig-height: 8
    embed-resources: true
    toc: true
    cache: false
    warning: false
---

```{r, include=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(reshape2)
library(RColorBrewer)
library(glue)

files_path <- fread(params$file_list, header = FALSE)
label_map <- list(
  "coverage-matrix.csv"              = "coverage_matrix",
  "estimate-cell-counts.csv"         = "estimate_cell_count",
  "methylation-matrix.csv"           = "methylation_matrix",
  "cutadapt_filtered_reads_plot.txt" = "reads",
  "picard_hsmetrics_table.txt"       = "hs"
)

for (path in files_path$V1) {
  filename <- basename(path)
  if (filename %in% names(label_map)) {
    var_name <- label_map[[filename]]
    assign(var_name, fread(path), envir = .GlobalEnv)
  }
}

read_counts_list <- list()

if (length(read_counts_list) > 0) {
  read_counts <- data.table::rbindlist(read_counts_list, fill = TRUE)
}

setnames(reads, old = "Reads passing filters", new = "Reads")
reads                            <- reads %>% mutate(Sample = gsub("(_R[12]|_[12])$", "", Sample)) %>% distinct(Sample, .keep_all = TRUE)
colnames(estimate_cell_count)[1] <- "cell_type"
panel                            <- fread(params$panel_data)
setkey(panel, chr, start, end)
setkey(methylation_matrix ,  chr, start, end)
input.data         <- foverlaps(methylation_matrix, panel, type = "within", nomatch = 0L)
input.coverage     <- foverlaps(coverage_matrix, panel, type = "within", nomatch = 0L)
input.data         <- input.data %>% select(-matches("chr|start|end|details|source"))
methylation_matrix <- methylation_matrix %>% select(-matches("chr|start|end"))
coverage_matrix    <- coverage_matrix %>% select(-matches("chr|start|end"))
input.coverage     <- input.coverage  %>% select(-matches("chr|start|end|details|source|i.start|i.end"))

```
## <sup>Panel Overview Table</sup>

```{r}
#| echo: false
#| label: panel-overview
#| results: asis

total_cpg_all       <- nrow(methylation_matrix)
total_cpg_panelfile <- nrow(panel)
cpg_after_filter    <- nrow(input.data)
num_samples         <- ncol(input.data)
coverage_threshold  <- 10

panel_summary <- data.frame(
  Metric = c(
    "Total CpG sites in methylation matrix (post-coverage filter)", 
    "CpG sites defined in panel regions",
    "CpG sites that overlap panel regions",
    "Number of samples analysed",
    glue("Coverage threshold (mincov = {coverage_threshold})")
  ),
  Value = format(
    c(total_cpg_all,
      total_cpg_panelfile,
      cpg_after_filter,
      num_samples,
      coverage_threshold),
    big.mark = ",")
)

cat('
<details style="margin-top:1rem;">
  <summary style="font-weight:bold; cursor:pointer;">
    What is the <code>methylation_matrix</code> and how was it generated?
  </summary>
  <div style="padding-left:10px; font-size:0.9rem;">
    <p>
      <code>methylation_matrix</code> is a sample × CpG table where each cell
      contains the fractional methylation (<em>β</em>-value) at that locus:
      <br><code>β = methylated_reads / (methylated_reads + unmethylated_reads)</code>.
    </p>
    <p><strong>Pipeline overview:</strong></p>
    <ol style="margin-left:1rem;">
      <li><strong>Adapter trimming</strong> – raw FASTQs processed with
          <code>Trim Galore!</code> (which uses <code>cutadapt</code>).</li>
      <li><strong>Alignment</strong> – trimmed reads aligned to the reference
          genome with <code>bwa-meth</code>; duplicates removed with <code>Picard</code>.</li>
      <li><strong>CpG extraction</strong> – per-CpG methylation counts generated
          by the aligner’s methyl-call utility.</li>
      <li><strong>Import to R</strong> – counts loaded with
          <code>methylKit::methRead()</code>; set
          <code>pipeline</code> and <code>assembly</code> to match your workflow.</li>
      <li><strong>Coverage filter</strong> – loci with coverage &lt;
          <strong>mincov = 10</strong> are discarded.
          <br><em>mincov</em> is the minimum number of reads required at a CpG
          to accept its β-value; lower coverage inflates variance and can
          distort downstream statistics.</li>
      <li><strong>Matrix export</strong> – the resulting
          <code>methylKit</code> object is converted to a matrix
          (<code>unite()</code> → <code>regionCounts()</code>)
          and saved as the <code>methylation_matrix</code>.</li>
    </ol>
    <p>
      Row count therefore equals the number of CpG loci that pass alignment,
      deduplication, and the <code>mincov</code> filter
      <em>before</em> any panel-based intersection.
    </p>
  </div>
</details>
')
```
## <sup>Methylation Distribution Plot for Panel</sup>

* This plot shows the distribution of DNA methylation levels across multiple samples. 
* The x-axis represents methylation values from 0 (unmethylated) to 1 (fully methylated), while the y-axis shows the density of CpG sites. 
* Each line corresponds to a sample, with color shading indicating overlapping densities. 
* The bimodal pattern, with peaks near 0 and 1, is typical for methylation data, highlighting unmethylated and highly methylated CpG regions. 
* This plot captures methylation variability across samples.

```{r}
#| echo: false
#| label: Density Plot of Methylation
#| code-fold: false
#| fig-cap: "**Figure 1**: Density Plot of Methylation for Samples."

methylation_data_long <- input.data %>%
  pivot_longer(cols = everything(), names_to = "Sample", values_to = "Methylation")

ggplot(methylation_data_long,
       aes(x = Methylation, fill = sub(".*_(S\\d+)$", "\\1", Sample),
           colour = sub(".*_(S\\d+)$", "\\1", Sample))) +
  geom_density(alpha = 0.25, linewidth = 0.6) +
  labs(title = "Density Plot of Methylation for Samples",
       x = "Methylation", 
       y = "Density",
       fill = "Sample", 
       colour = "Sample") + 
  theme(legend.position = "right") 
```
## <sup>Estimate cell counts</sup>

* The plot shows data density with labels for Q1, median, and Q3.

```{r}  
#| echo: false
#| code-fold: false
#| label: Violin Plot
#| fig-cap: "**Figure 2**: The violin plot displays the distribution of estimated cell counts across cell types."

getPalette <- colorRampPalette(brewer.pal(8, "Set2"))
data <- estimate_cell_count %>%
  pivot_longer(cols = -cell_type, names_to = "Sample", values_to = "Value") %>%
  rename(CellType = cell_type)

iqr_data <- data %>%
  group_by(CellType) %>%
  summarize(
    Q1 = quantile(Value, 0.25),
    Median = median(Value),
    Q3 = quantile(Value, 0.75))
violin_plot <- ggplot(data, aes(x = CellType, y = Value, fill = CellType)) +
  geom_violin(alpha = 0.25, 
   scale    = "width",
   trim     = FALSE,
   position = position_dodge(width = 0.5)) + 
  scale_fill_manual(values = getPalette(length(unique(data$CellType)))) +  
  geom_text(data = iqr_data, aes(x = CellType, y = Q3, label = paste0("Q3: ", round(Q3, 2))), 
            position = position_nudge(x = 0.2), size = 3) + 
  geom_text(data = iqr_data, aes(x = CellType, y = Median, label = paste0("Median: ", round(Median, 2))), 
            position = position_nudge(x = 0.2), size = 3) +  
  geom_text(data = iqr_data, aes(x = CellType, y = Q1, label = paste0("Q1: ", round(Q1, 2))), 
            position = position_nudge(x = 0.2), size = 3) +  
  theme_minimal(base_size = 15) +  
  labs(
    title = "Violin Plot: Distribution of Cell Types",
    x = "Cell Type",
    y = "Value"
  ) +
  theme(
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, hjust = 0.5),
    legend.position = "none"  
  )
print(violin_plot)
```
```{r}
#| echo: false 
#| label: boxplot
#| code-fold: false
#| fig-cap: "**Figure 3**: The boxplot displays the distribution of estimated cell counts across cell types."

boxplot_plot <- ggplot(data, aes(x = CellType, y = Value, fill = CellType)) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.8), outlier.shape = NA) + 
  scale_fill_manual(values = getPalette(length(unique(data$CellType)))) +  
  theme_minimal(base_size = 15) +  
  labs(
    title = "Boxplot: Distribution of Cell Types",
    x = "Cell Type",
    y = "Value"
  ) +
  theme(
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8))
print(boxplot_plot)
```
## <sup>Depth Distribution</sup>

* This plot illustrates the distribution of depth for each sample, with Samples represented along the x-axis and depth values on the y-axis.
* Side-by-side view: Two boxes per sample—Panel and Total—compare on-target depth to genome-wide depth at a glance.
* Sorted by quality: Samples are arranged by median Panel depth, putting the best on the left and weak outliers on the right.
* Quick red-flags: Panel boxes should be higher and tighter; lower or wider ones signal poor capture or uneven depth.

```{r}
#| echo: false
#| label: Depth Distribution Violin Plot
#| code-fold: false
#| fig-cap: "**Figure 4**:  Comparison of Depth per Sample Between Panel CpG sites and Total CpG sites."

depth_long <- bind_rows(
  input.coverage   %>% 
    mutate(Set = "Panel") %>%          
    pivot_longer(cols = -Set,
                 names_to = "Sample",
                 values_to = "Depth"),
  coverage_matrix %>% 
    mutate(Set = "Total") %>% 
    pivot_longer(cols = -Set,
                 names_to = "Sample",
                 values_to = "Depth")
)

depth_long <- depth_long %>%
  mutate(SampleLabel = sub(".*_(S\\d+)", "\\1", Sample))

sample_order <- depth_long %>% 
  group_by(SampleLabel) %>% 
  summarise(mean_cov = mean(Depth, na.rm = TRUE), .groups = "drop") %>% 
  arrange(desc(mean_cov)) %>% 
  pull(SampleLabel)

depth_long$SampleLabel <- factor(depth_long$SampleLabel,
                                    levels = sample_order)

ymax <- max(depth_long$Depth, na.rm = TRUE)

ggplot(
  depth_long,
  aes(x = SampleLabel, y = Depth, fill = Set)
) +
  geom_violin(
    trim     = FALSE,
    scale    = "width",
    width    = .8,
    alpha    = .70,
    position = position_dodge(width = .8)
  ) +
  stat_summary(
    fun      = median,
    geom     = "point",
    size     = .5,
    colour   = "black",
    position = position_dodge(width = .8)
  ) +
  scale_y_continuous(
    limits = c(0, ymax),
    expand = expansion(mult = c(0, .05))
  ) +
  scale_fill_manual(values = c(Panel = "#2c7bb6", Total = "#d7191c")) +
  labs(
    title = "Depth per Sample: Panel vs Total",
    x     = "Sample",
    y     = "Depth",
    fill  = "Set"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x        = element_text(angle = 90, vjust = 0.5, hjust = 1),
    panel.grid.major.x = element_blank()
  ) 
```
## <sup>PCA for Methylation</sup>

```{r, include=FALSE}
#| echo: false
#| label: PCA Scaled functions
#| code-fold: false
#| error: true

impute.matrix <- function(x, margin = 1, fun = function(x) mean(x, na.rm = TRUE)) {
    if (margin == 2) x <- t(x)
    idx <- which(is.na(x) | !is.finite(x), arr.ind = TRUE)
    if (length(idx) > 0) {
        na.idx <- unique(idx[,"row"])
        v <- apply(x[na.idx, , drop = FALSE], 1, fun)
        v[which(is.na(v))] <- fun(v)
        x[idx] <- v[match(idx[,"row"], na.idx)]
        stopifnot(all(!is.na(x)))
    }
    if (margin == 2) x <- t(x)
    x
}
# Function to perform PCA and generate scree plot and scatter plots
generate_pca_plots <- function(data, scale_data=TRUE, dataset_name = "Scaled Data") { 
  # Perform PCA
  pca_result <- prcomp(t(data), scale. = scale_data)
  
  # Calculate explained variance
  eigenvalues <- (pca_result$sdev)^2
  explained_variance <- eigenvalues / sum(eigenvalues)
  
  # Create Scree Plot
  scree_plot <- ggplot(data = data.frame(PC = 1:length(explained_variance), Variance = explained_variance), 
                       aes(x = PC, y = Variance)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(title = paste("Scree Plot -", dataset_name), x = "Principal Components", y = "Explained Variance") +
    theme_minimal()
  
  # Create PCA scores data frame
  pca_scores <- as.data.frame(pca_result$x)
  rownames(pca_scores) <- sub("\\.", "-", gsub("^X", "",rownames(pca_scores)))
  pca_scores$Sample <- rownames(pca_scores)
  
  # Merge with reads data
  pca_scores <- merge(pca_scores, reads, by.x = "Sample", by.y = "Sample", all.x = TRUE)
  
  # Create PCA scatter plots
  pcs <- list(c("PC1","PC2"),c("PC1","PC3"),c("PC2","PC3"),c("PC3","PC4"))
  scatter_plots <- lapply(pcs, function(pcs) {
      ggplot(pca_scores, aes_string(x = pcs[1], y = pcs[2], color = "Reads")) +
        geom_point() +
        geom_text(aes(label = sub(".*_(S\\d+)", "\\1", Sample, "")), vjust = -0.5, hjust = 0.5, size = 3) + 
        theme_minimal() +
        labs(title = paste("PCA Scatterplot:", pcs[1] , "vs", pcs[2], dataset_name), color = "Reads") +
        scale_color_gradient(low = "blue", high = "red")
  })
  
  # Return scree plot and scatter plots
  return(list(scree_plot = scree_plot, scatter_plots = scatter_plots))
}
```
```{r}
#| echo: false
#| label: PCA Scaled
#| code-fold: false
#| error: true

# Generate plots for Scaled data
  meth <- impute.matrix(data.frame(input.data), 2)
  # Generate PCA plots for scaled data
  scaled_plots <- generate_pca_plots(meth, scale_data = TRUE, dataset_name = "Scaled Data")
```
```{r}
#| echo: false
#| label: Scree Plot Scaled
#| error: true
#| fig-cap: "**Figure 5**: Scree Plot for Scaled Data. This plot shows the proportion of variance explained by the principal components for the scaled data. The first few components explain most of the variance."
print(scaled_plots$scree_plot)
```
```{r}
#| echo: false
#| label: Scatter Plot PC1 vs PC2 Scaled
#| error: true
#| fig-cap: "**Figure 6**: PCA Scatterplot of PC1 vs PC2 for Scaled Data. Each point represents a sample, color-coded based on the number of reads for the scaled dataset."
print(scaled_plots$scatter_plots[[1]])
```
```{r}
#| echo: false
#| label: Scatter Plot PC1 vs PC3 Scaled
#| error: true
#| fig-cap: "**Figure 7**: PCA Scatterplot of PC1 vs PC3 for Scaled Data. This plot shows the relationship between PC1 and PC3 for the scaled data."
print(scaled_plots$scatter_plots[[2]])
```
```{r}
#| echo: false
#| label: Scatter Plot PC2 vs PC3 Scaled
#| error: true
#| fig-cap: "**Figure 8**: PCA Scatterplot of PC2 vs PC3 for Scaled Data. This plot shows the relationship between PC2 and PC3 for the scaled data, with each point color-coded by read count."
print(scaled_plots$scatter_plots[[3]])
```
```{r}
#| echo: false
#| label: Scatter Plot PC3 vs PC4 Scaled
#| error: true
#| fig-cap: "**Figure 9**: PCA Scatterplot of PC3 vs PC4 for Scaled Data. This plot shows the relationship between PC3 and PC4 for the scaled data, with each point color-coded by read count."
print(scaled_plots$scatter_plots[[4]])
```
## <sup>Correlation for Methylation</sup>
```{r}
#| echo: false
#| label: Correlation for Methylation Raw/Scaled
#| error: true
#| code-fold: false

generate_correlation_plots <- function(data) {  
  # Calculate correlation matrices
  cor_matrix_no_standardize <- cor(data, use = "pairwise.complete.obs")
  meth_data_scaled_by_cpg <- t(scale(t(data)))  # Standardize CpG sites row-wise
  cor_matrix_standardized <- cor(meth_data_scaled_by_cpg, use = "pairwise.complete.obs")
  
  # Melt correlation matrices for plotting
  cor_melt_no_standardize <- melt(cor_matrix_no_standardize)
  cor_melt_standardized <- melt(cor_matrix_standardized)
  
  # Plot density of correlations without standardization
  p1 <- ggplot(cor_melt_no_standardize, aes(x = value)) +
    geom_density(fill = "blue", alpha = 0.5) +
    labs(title = paste("Density Plot Without Standardizing CpG Sites"), 
         x = "Correlation", y = "Density") +
    theme_minimal()
  
  # Plot density of correlations with standardization
  p2 <- ggplot(cor_melt_standardized, aes(x = value)) +
    geom_density(fill = "red", alpha = 0.5) +
    labs(title = paste("Density Plot With Standardizing CpG Sites"), 
         x = "Correlation", y = "Density") +
    theme_minimal()
  
  # Print the plots
   return(list(p1 = p1, p2 = p2))
} 
  # Generate correlation plots for full dataset
 plots <- generate_correlation_plots(input.data)
```
* This plot shows the distribution of pairwise correlations between samples without standardizing the CpG sites. 
* The plot provides an overview of the raw correlation data.

```{r}
#| echo: false
#| label: Density Plot Without Standardizing
#| error: true
#| fig-cap: "**Figure 10**: Density Plot Without Standardizing CpG Sites."

# Print the density plot for correlations without standardization
print(plots$p1)
```
* This plot shows the distribution of pairwise correlations between samples after standardizing the CpG sites. 
* Standardization normalizes CpG sites across samples, offering a clearer view of relative variations in the data.

```{r}
#| echo: false
#| label: Density Plot With Standardizing
#| error: true
#| fig-cap: "**Figure 11**: Density Plot With Standardizing CpG Sites."
print(plots$p2)
```
## <sup>Total Read Count per Sample</sup>

* Read counts: This plot tracks the total number of mapped reads overlapping CpG loci for each library, directly calculated from sorted BAM files using samtools. 
* Panel vs. total at a glance: The plot also highlights, for each sample, how many reads map to the whole methylome (green bar: total) and to the targeted panel regions (blue bar: panel). 
* This enables quick visual detection of libraries with sufficient total sequencing depth but suboptimal panel capture.

```{r}
#| echo: false
#| label: total-read-count 
#| error: true
#| fig-cap: "**Figure 12**: Horizontal bar plot for each sample, showing the **total number of mapped reads** overlapping CpG loci in the whole methylome (green) and in the targeted panel regions (blue)."

long_counts <- pivot_longer(read_counts, 
                            cols = c(total_reads, panel_reads), 
                            names_to = "Type", values_to = "Reads")

long_counts$Type <- factor(long_counts$Type, levels = c("panel_reads", "total_reads"))

long_counts$SampleID <- factor(long_counts$SampleID, 
    levels = read_counts$SampleID[order(-read_counts$total_reads)])

custom_colors <- c("total_reads" = "#38B000", "panel_reads" = "#0077B6")

ggplot(long_counts, aes(x = Reads, y = SampleID, fill = Type, label = Reads)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "grey95", linewidth = 0.18) +
  geom_text(position = position_dodge(width = 0.8), hjust = -0.10, size = 3.5, fontface = "plain", family = "Times New Roman") +
  scale_fill_manual(
    values = custom_colors,
    labels = c("Panel Reads", "Total Reads")
  ) +
  labs(
    title = "Read Counts per Sample",
    subtitle = "Each sample: Total & Panel reads",
    x = "Read Count", y = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 15, base_family = "Times New Roman") +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12, family = "Times New Roman"),
    plot.title = element_text(face = "bold", size = 18, hjust = 0, family = "Times New Roman"),
    plot.subtitle = element_text(size = 13, hjust = 0, family = "Times New Roman"),
    axis.title.x = element_text(face = "bold", size = 14, margin = margin(t = 10), family = "Times New Roman"),
    axis.text.x = element_text(face = "plain", size = 12, family = "Times New Roman"),
    axis.text.y = element_text(face = "plain", size = 12, family = "Times New Roman"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = "grey90", linewidth = 0.2)
  ) +
  xlim(0, max(long_counts$Reads) * 1.13)
```
## <sup>Panel CpG Coverage Completeness per Sample</sup>

* This metric evaluates how completely each sample covers the predefined CpG panel. 
* Completeness is calculated as the proportion (%) of targeted CpG sites from the panel that were successfully detected, defined as having at least one mapped read.
* Bars: Represent the count of successfully covered CpG sites per sample.
* Percentage labels: Directly reflect completeness relative to the total panel size.

```{r}
#| echo: false
#| label: total-read-count panel  
#| error: true
#| fig-cap: "**Figure 13**: CpG Coverage Completeness per Sample"

total_panel_sites <- nrow(input.coverage)   
covered_panel_sites <- colSums(input.coverage > 0, na.rm = TRUE)

coverage_df <- tibble(
  Sample              = names(covered_panel_sites),
  Covered_Panel_Sites = covered_panel_sites
) |>
  mutate(
    Covered_Pct = 100 * Covered_Panel_Sites / total_panel_sites
  ) |>
  arrange(desc(Covered_Panel_Sites))                    

ggplot(coverage_df,
       aes(x = reorder(Sample, Covered_Panel_Sites),
           y = Covered_Panel_Sites)) +
  geom_col(fill = "#4E79A7", width = 0.75,
           colour = "grey20", linewidth = 0.15) +
  geom_text(aes(label = sprintf("%.1f%%", Covered_Pct)),
            hjust = -0.15, size = 3) +
  coord_flip(clip = "off") +
  scale_y_continuous(
    limits = c(0, total_panel_sites * 1.05),
    breaks = scales::pretty_breaks(5)
  ) +
  labs(
    title = "Panel CpG Sites Covered per Sample",
    subtitle = glue::glue("{total_panel_sites} total panel CpG sites"),
    x = "Sample",
    y = "CpG Sites"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.major.y = element_blank()
  )
```
## <sup>Sequencing Coverage Uniformity</sup>

* This graph visually evaluates the uniformity of sequencing coverage for each sample across the targeted regions.
* Samples located close to the red diagonal line (where Mean ≈ Median) have a more homogeneous coverage distribution, which is the desired outcome.
* Samples that fall significantly below the line have a skewed distribution. 
* In these cases, the average (mean) coverage is artificially inflated by a few "hotspot" regions with extremely high depth, while the bulk of the regions have coverage closer to the lower median value.
* The position on the axes also indicates the overall sequencing depth—samples in the top-right corner have the highest coverage.

```{r}
#| echo: false
#| label: Assessment of Sequencing Coverage Uniformity
#| error: true
#| fig-cap: "**Figure 14**:This scatter plot compares the mean and median target coverage for each sample to evaluate the homogeneity of the sequencing data. The dashed diagonal line represents a perfect 1:1 ratio, indicating a completely uniform coverage distribution."

ggplot(hs,
       aes(`Mean target coverage`,
           `Median target coverage`,
           colour = Sample)) +
  geom_point(size = 2.2, alpha = .85,
             position = position_jitter(width = .02, height = .04)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "red") +
  scale_x_continuous(limits = c(0, max(hs$`Mean target coverage`))) +
  scale_y_continuous(limits = c(0, max(hs$`Mean target coverage`))) +
  scale_colour_discrete(labels = function(x) sub(".*_(S\\d+)", "\\1", x)) +
  labs(title  = "Mean vs Median Target Coverage",
       x      = "Mean target coverage",
       y      = "Median target coverage",
       colour = "Sample") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "right")
```
