---
title: "DNA Methylation: Methyl-seq Panel QC Report"
author: "."
date: "`r Sys.Date()`"
encoding: UTF-8 
params:
  meth: NULL
  coverage: NULL
  cell_counts: NULL
  filtered_reads: NULL
  raw_reads: NULL
  hs: NULL
  panel: NULL
format:
  html:
    fig-cap-location: bottom
    df-print: paged
    fig-width: 10
    fig-height: 8
    embed-resources: true
    toc: true
    cache: false
    warning: false
---

```{r, include=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(reshape2)
library(RColorBrewer)
library(glue)

methylation_matrix <- fread(params$meth)
coverage_matrix <- fread(params$coverage)
cell.counts <- fread(params$cell_counts)
filtered.reads <- fread(params$filtered_reads)
raw.reads <- fread(params$raw_reads)
hs <- fread(params$hs)
panel <- fread(params$panel)

setnames(filtered.reads, old = "Reads passing filters", new = "Reads")
colnames(filtered.reads) <- c("sample","reads")
filtered.reads$sample <- sub("(_R[12]|_[12])$", "", filtered.reads$sample)
filtered.reads <- unique(filtered.reads)

colnames(cell.counts)[1] <- "cell_type"

setkey(panel, chr, start, end)
setkey(methylation_matrix ,  chr, start, end)
input.data         <- foverlaps(methylation_matrix, panel, type = "within", nomatch = 0L)
input.coverage     <- foverlaps(coverage_matrix, panel, type = "within", nomatch = 0L)
input.data         <- input.data %>% select(-matches("chr|start|end|details|source"))
methylation_matrix <- methylation_matrix %>% select(-matches("chr|start|end"))
coverage_matrix    <- coverage_matrix %>% select(-matches("chr|start|end"))
input.coverage     <- input.coverage  %>% select(-matches("chr|start|end|details|source|i.start|i.end"))
```
* <em>All metrics and plots in this report refer to the predefined CpG panel unless explicitly noted otherwise.</em>

## <sup>Overview Table</sup>

```{r}
#| echo: false
#| label: panel-overview
#| results: asis

# Numbers

total_cpg_all       <- nrow(methylation_matrix)
total_cpg_panelfile <- nrow(panel)
cpg_after_filter    <- nrow(input.data)
num_samples         <- ncol(input.data)
coverage_threshold  <- 10

#Descriptions

metrics <- c(
  "Covered CpG sites",
  "Targeted CpG sites",
  "Targeted and covered CpG sites",
  "Number of samples analysed",
  "Coverage threshold (mincov)"
)

descriptions <- c(
  "The number of CpG sites in the human genome for which at least 50% of the samples each have at least 'mincov' reads covering the site.",
  "The number of CpG sites inside panel regions",
  "The number of covered CpG sites inside panel regions",
  "",
  ""
)

values <- format(
  c(total_cpg_all,
    total_cpg_panelfile,
    cpg_after_filter,
    num_samples,
    coverage_threshold),
  big.mark = ","
)

row_html <- vapply(seq_along(metrics), function(i) {
  glue('  <tr>
    <td style="padding:6px 10px;">
      {metrics[i]}<br>
      <span class="subdesc">{descriptions[i]}</span>
    </td>
    <td style="padding:6px 10px; text-align:right;">{values[i]}</td>
  </tr>')
}, character(1))

# CSS codes

css <- '
<style>
  /* Table size & typography */
  table.metric-table {
    width:90%;
    border-collapse:collapse;
    font-size:0.95rem;
    margin:1rem auto;
  }

  /* Subtle gray borders */
  .metric-table,
  .metric-table th,
  .metric-table td {
    border:1px solid #d8d8d8;
  }

  /* Header row – light grey background, darker text */
  .metric-table th {
    background:#f2f2f2;
    color:#000000;
    text-align:left;
    padding:6px 10px;
    font-weight:600;
  }

  /* Data cells – white + gentle zebra striping */
  .metric-table td {
    background:#ffffff;
    color:#333333;
    padding:6px 10px;
  }
  .metric-table tr:nth-child(even) td { background:#fafafa; }

  /* Description under each metric – darker & semi-bold */
  .metric-table .subdesc {
    font-style:italic;
    font-size:0.85em;
    color:#4a4a4a;
    font-weight:600;
  }
</style>
'

#HTML out

html_out <- glue('{css}
<table class="metric-table">
  <thead>
    <tr><th>Metric</th><th style="text-align:right;">Value</th></tr>
  </thead>
  <tbody>
{paste(row_html, collapse = "\n")}
  </tbody>
</table>
')

cat(html_out)
```
## <sup>Methylation Distribution Plot</sup>

* This plot shows the distribution of DNA methylation levels across multiple samples. 
* The x-axis represents methylation values from 0 (unmethylated) to 1 (fully methylated),
* while the y-axis shows the density of CpG sites. 
* Each line corresponds to a sample, with color shading indicating overlapping densities. 
* The bimodal pattern, with peaks near 0 and 1, is typical for methylation data,
* highlighting unmethylated and highly methylated CpG regions. 
* This plot captures methylation variability across samples.

```{r}
#| echo: false
#| label: Density Plot of Methylation
#| code-fold: false
#| fig-cap: "**Figure 1**: Density Plot of Methylation for Samples."

methylation_data_long <- input.data %>%
  pivot_longer(cols = everything(), names_to = "Sample", values_to = "Methylation")

ggplot(methylation_data_long,
       aes(x = Methylation, fill = sub(".*_(S\\d+)$", "\\1", Sample),
           colour = sub(".*_(S\\d+)$", "\\1", Sample))) +
  geom_density(alpha = 0.25, linewidth = 0.6) +
  labs(title = "Density Plot of Methylation for Samples",
       x = "Methylation", 
       y = "Density",
       fill = "Sample", 
       colour = "Sample") + 
  theme(legend.position = "right") 
```
## <sup>Estimate cell counts</sup>

* The plot shows data density with labels for Q1, median, and Q3.
* These cell-type proportions are inferred from genome-wide methylation patterns
* and do not rely on the targeted CpG panel. 
* In other words, the estimates reflect the overall cellular composition of each sample,
* independent of panel capture efficiency or on-target depth.

```{r}  
#| echo: false
#| code-fold: false
#| label: Violin Plot
#| fig-cap: "**Figure 2**: The violin plot displays the distribution of estimated cell counts across cell types."

tryCatch({
  
getPalette <- colorRampPalette(brewer.pal(8, "Set2"))
data <- cell.counts %>%
  pivot_longer(cols = -cell_type, names_to = "Sample", values_to = "Value") %>%
  rename(CellType = cell_type)

iqr_data <- data %>%
  group_by(CellType) %>%
  summarize(
    Q1 = quantile(Value, 0.25),
    Median = median(Value),
    Q3 = quantile(Value, 0.75))
violin_plot <- ggplot(data, aes(x = CellType, y = Value, fill = CellType)) +
  geom_violin(alpha = 0.25, 
   scale    = "width",
   trim     = FALSE,
   position = position_dodge(width = 0.5)) + 
  scale_fill_manual(values = getPalette(length(unique(data$CellType)))) +  
  geom_text(data = iqr_data, aes(x = CellType, y = Q3, label = paste0("Q3: ", round(Q3, 2))), 
            position = position_nudge(x = 0.2), size = 3) + 
  geom_text(data = iqr_data, aes(x = CellType, y = Median, label = paste0("Median: ", round(Median, 2))), 
            position = position_nudge(x = 0.2), size = 3) +  
  geom_text(data = iqr_data, aes(x = CellType, y = Q1, label = paste0("Q1: ", round(Q1, 2))), 
            position = position_nudge(x = 0.2), size = 3) +  
  theme_minimal(base_size = 15) +  
  labs(
    title = "Violin Plot: Distribution of Cell Types",
    x = "Cell Type",
    y = "Value"
  ) +
  theme(
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, hjust = 0.5),
    legend.position = "none"  
  )
print(violin_plot)

}, error=function(msg) {
  print(msg)
})
```
```{r}
#| echo: false 
#| label: boxplot
#| code-fold: false
#| fig-cap: "**Figure 3**: The boxplot displays the distribution of estimated cell counts across cell types."

tryCatch({
  
boxplot_plot <- ggplot(data, aes(x = CellType, y = Value, fill = CellType)) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.8), outlier.shape = NA) + 
  scale_fill_manual(values = getPalette(length(unique(data$CellType)))) +  
  theme_minimal(base_size = 15) +  
  labs(
    title = "Boxplot: Distribution of Cell Types",
    x = "Cell Type",
    y = "Value"
  ) +
  theme(
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8))
print(boxplot_plot)
}, error=function(msg) {
  print(msg)
})
```
## <sup>Depth Distribution</sup>

* This plot illustrates the distribution of depth for each sample,
* with Samples represented along the x-axis and depth values on the y-axis.
* Side-by-side view: Two boxes per sample—Panel and Total—compare
* on-target depth to genome-wide depth at a glance.
* Sorted by quality: Samples are arranged by median Panel depth,
* putting the best on the left and weak outliers on the right.
* Quick red-flags: Panel boxes should be higher and tighter;
* lower or wider ones signal poor capture or uneven depth.

```{r}
#| echo: false
#| label: Depth Distribution Violin Plot
#| code-fold: false
#| fig-cap: "**Figure 4**:  Comparison of Depth per Sample Between Panel CpG sites and Total CpG sites."

depth_long <- bind_rows(
  input.coverage   %>% 
    mutate(Set = "Panel") %>%          
    pivot_longer(cols = -Set,
                 names_to = "Sample",
                 values_to = "Depth"),
  coverage_matrix %>% 
    mutate(Set = "Total") %>% 
    pivot_longer(cols = -Set,
                 names_to = "Sample",
                 values_to = "Depth")
)

depth_long <- depth_long %>%
  mutate(SampleLabel = sub(".*_(S\\d+)", "\\1", Sample))

sample_order <- depth_long %>% 
  group_by(SampleLabel) %>% 
  summarise(mean_cov = mean(Depth, na.rm = TRUE), .groups = "drop") %>% 
  arrange(desc(mean_cov)) %>% 
  pull(SampleLabel)

depth_long$SampleLabel <- factor(depth_long$SampleLabel,
                                    levels = sample_order)

ymax <- max(depth_long$Depth, na.rm = TRUE)

ggplot(
  depth_long,
  aes(x = SampleLabel, y = Depth, fill = Set)
) +
  geom_violin(
    trim     = FALSE,
    scale    = "width",
    width    = .8,
    alpha    = .70,
    position = position_dodge(width = .8)
  ) +
  stat_summary(
    fun      = median,
    geom     = "point",
    size     = .5,
    colour   = "black",
    position = position_dodge(width = .8)
  ) +
  scale_y_continuous(
    limits = c(0, ymax),
    expand = expansion(mult = c(0, .05))
  ) +
  scale_fill_manual(values = c(Panel = "#2c7bb6", Total = "#d7191c")) +
  labs(
    title = "Depth per Sample: Panel vs Total",
    x     = "Sample",
    y     = "Depth",
    fill  = "Set"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x        = element_text(angle = 90, vjust = 0.5, hjust = 1),
    panel.grid.major.x = element_blank()
  ) 
```
## <sup>Principal Component Analysis (PCA)</sup>

```{r, include=FALSE}
#| echo: false
#| label: PCA Scaled functions
#| code-fold: false
#| error: true

impute.matrix <- function(x, margin = 1, fun = function(x) mean(x, na.rm = TRUE)) {
    if (margin == 2) x <- t(x)
    idx <- which(is.na(x) | !is.finite(x), arr.ind = TRUE)
    if (length(idx) > 0) {
        na.idx <- unique(idx[,"row"])
        v <- apply(x[na.idx, , drop = FALSE], 1, fun)
        v[which(is.na(v))] <- fun(v)
        x[idx] <- v[match(idx[,"row"], na.idx)]
        stopifnot(all(!is.na(x)))
    }
    if (margin == 2) x <- t(x)
    x
}
# Function to perform PCA and generate scree plot and scatter plots

generate_pca_plots <- function(data, scale_data = TRUE, dataset_name = "Scaled Data") {
  ## Remove CpG loci with zero or undefined variance (row-wise QC)
  var_row <- apply(data, 1, var, na.rm = TRUE)
  keep    <- !is.na(var_row) & var_row > 1e-25      # ε = 1e-25 tolerance
  if (!all(keep)) {
    dropped <- sum(!keep)
    warning(sprintf(
      "%d CpG locus/loci with zero or undefined variance removed before PCA",
      dropped))
    data <- data[keep, , drop = FALSE]
  }

  pca_result <- prcomp(t(data), scale. = scale_data)
  ## scree plot + score scatter plots
  eigenvalues        <- pca_result$sdev^2
  explained_variance <- eigenvalues / sum(eigenvalues)

  scree_plot <- ggplot(
      data.frame(PC = seq_along(explained_variance),
                 Variance = explained_variance),
      aes(PC, Variance)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(title = paste("Scree Plot -", dataset_name),
         x = "Principal Components", y = "Explained Variance") +
    theme_minimal()

  # Prepare score matrix
  pca_scores           <- as.data.frame(pca_result$x)
  rownames(pca_scores) <- sub("\\.", "-", gsub("^X", "", rownames(pca_scores)))
  pca_scores$sample    <- rownames(pca_scores)
  pca_scores           <- merge(pca_scores, filtered.reads,
                                by = "sample", all.x = TRUE)

  # Define PC pairs to plot
  pc_pairs <- list(c("PC1","PC2"), c("PC1","PC3"),
                   c("PC2","PC3"), c("PC3","PC4"))

  scatter_plots <- lapply(pc_pairs, function(ax) {
    ggplot(pca_scores,
           aes_string(ax[1], ax[2], color = "reads")) +
      geom_point() +
      geom_text(aes(label = sub(".*_(S\\d+)", "\\1", sample)),
                vjust = -0.5, hjust = 0.5, size = 3) +
      theme_minimal() +
      labs(title = paste("PCA:", ax[1], "vs", ax[2], "-", dataset_name),
           color = "reads") +                       
      scale_color_gradient(low = "blue", high = "red")
  })

  list(scree_plot = scree_plot, scatter_plots = scatter_plots)
}
```
```{r}
#| echo: false
#| label: PCA Scaled
#| code-fold: false
#| error: true

# Generate plots for Scaled data
  meth <- impute.matrix(data.frame(input.data), 2)
  # Generate PCA plots for scaled data
  scaled_plots <- generate_pca_plots(meth, scale_data = TRUE, dataset_name = "Scaled Data")
```
```{r}
#| echo: false
#| label: Scree Plot Scaled
#| error: true
#| fig-cap: "**Figure 5**: Scree Plot for Scaled Data. This plot shows the proportion of variance explained by the principal components for the scaled data. The first few components explain most of the variance."
print(scaled_plots$scree_plot)
```
```{r}
#| echo: false
#| label: Scatter Plot PC1 vs PC2 Scaled
#| error: true
#| fig-cap: "**Figure 6**: PCA Scatterplot of PC1 vs PC2 for Scaled Data. Each point represents a sample, color-coded based on the number of reads for the scaled dataset."
print(scaled_plots$scatter_plots[[1]])
```
```{r}
#| echo: false
#| label: Scatter Plot PC1 vs PC3 Scaled
#| error: true
#| fig-cap: "**Figure 7**: PCA Scatterplot of PC1 vs PC3 for Scaled Data. This plot shows the relationship between PC1 and PC3 for the scaled data."
print(scaled_plots$scatter_plots[[2]])
```
```{r}
#| echo: false
#| label: Scatter Plot PC2 vs PC3 Scaled
#| error: true
#| fig-cap: "**Figure 8**: PCA Scatterplot of PC2 vs PC3 for Scaled Data. This plot shows the relationship between PC2 and PC3 for the scaled data, with each point color-coded by read count."
print(scaled_plots$scatter_plots[[3]])
```
```{r}
#| echo: false
#| label: Scatter Plot PC3 vs PC4 Scaled
#| error: true
#| fig-cap: "**Figure 9**: PCA Scatterplot of PC3 vs PC4 for Scaled Data. This plot shows the relationship between PC3 and PC4 for the scaled data, with each point color-coded by read count."
print(scaled_plots$scatter_plots[[4]])
```
## <sup>Sample-wise Correlation</sup>
```{r}
#| echo: false
#| label: Correlation for Methylation Raw/Scaled
#| error: true
#| code-fold: false

generate_correlation_plots <- function(data) {  
  # Calculate correlation matrices
  cor_matrix_no_standardize <- cor(data, use = "pairwise.complete.obs")
  meth_data_scaled_by_cpg <- t(scale(t(data)))  # Standardize CpG sites row-wise
  cor_matrix_standardized <- cor(meth_data_scaled_by_cpg, use = "pairwise.complete.obs")
  
  # Melt correlation matrices for plotting
  cor_melt_no_standardize <- melt(cor_matrix_no_standardize)
  cor_melt_standardized <- melt(cor_matrix_standardized)
  
  # Plot density of correlations without standardization
  p1 <- ggplot(cor_melt_no_standardize, aes(x = value)) +
    geom_density(fill = "blue", alpha = 0.5) +
    labs(title = paste("Density Plot Without Standardizing CpG Sites"), 
         x = "Correlation", y = "Density") +
    theme_minimal()
  
  # Plot density of correlations with standardization
  p2 <- ggplot(cor_melt_standardized, aes(x = value)) +
    geom_density(fill = "red", alpha = 0.5) +
    labs(title = paste("Density Plot With Standardizing CpG Sites"), 
         x = "Correlation", y = "Density") +
    theme_minimal()
  
  # Print the plots
   return(list(p1 = p1, p2 = p2))
} 
  # Generate correlation plots for full dataset
 plots <- generate_correlation_plots(input.data)
```
* This plot shows the distribution of pairwise correlations between samples without standardizing the CpG sites. 
* The plot provides an overview of the raw correlation data.

```{r}
#| echo: false
#| label: Density Plot Without Standardizing
#| error: true
#| fig-cap: "**Figure 10**: Density Plot Without Standardizing CpG Sites."

# Print the density plot for correlations without standardization
print(plots$p1)
```
* This plot shows the distribution of pairwise correlations between samples after standardizing the CpG sites. 
* Standardization normalizes CpG sites across samples, offering a clearer view of relative variations in the data.

```{r}
#| echo: false
#| label: Density Plot With Standardizing
#| error: true
#| fig-cap: "**Figure 11**: Density Plot With Standardizing CpG Sites."
print(plots$p2)
```
## <sup>Read Count per Sample</sup>

* Read counts: This plot tracks the total number of mapped reads overlapping CpG loci for each library,
* directly calculated from sorted BAM files using samtools. 
* Panel vs. Total at a glance: The plot also highlights, for each sample,
* how many reads map to the whole methylome (green bar: total) and to the targeted panel regions (blue bar: panel). 
* This enables quick visual detection of libraries with sufficient total sequencing depth
* but suboptimal panel capture.

```{r}
#| echo: false
#| label: total-read-count 
#| error: true
#| fig-cap: "**Figure 12**: Horizontal bar plot for each sample, showing the **total number of mapped reads** overlapping CpG loci in the whole methylome (green) and in the targeted panel regions (blue)."

long_counts <- pivot_longer(raw.reads, 
                            cols = c(total_reads, panel_reads), 
                            names_to = "Type", values_to = "Reads")

long_counts$Type <- factor(long_counts$Type, levels = c("panel_reads", "total_reads"))

long_counts$SampleID <- factor(long_counts$SampleID, 
    levels = raw.reads$SampleID[order(-raw.reads$total_reads)])

custom_colors <- c("total_reads" = "#38B000", "panel_reads" = "#0077B6")

ggplot(long_counts, aes(x = Reads, y = SampleID, fill = Type, label = Reads)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "grey95", linewidth = 0.18) +
  geom_text(position = position_dodge(width = 0.8), hjust = -0.10, size = 3.5, fontface = "plain", family = "Times New Roman") +
  scale_fill_manual(
    values = custom_colors,
    labels = c("Panel Reads", "Total Reads")
  ) +
  labs(
    title = "Read Counts per Sample",
    subtitle = "Each sample: Total & Panel reads",
    x = "Read Count", y = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 15, base_family = "Times New Roman") +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12, family = "Times New Roman"),
    plot.title = element_text(face = "bold", size = 18, hjust = 0, family = "Times New Roman"),
    plot.subtitle = element_text(size = 13, hjust = 0, family = "Times New Roman"),
    axis.title.x = element_text(face = "bold", size = 14, margin = margin(t = 10), family = "Times New Roman"),
    axis.text.x = element_text(face = "plain", size = 12, family = "Times New Roman"),
    axis.text.y = element_text(face = "plain", size = 12, family = "Times New Roman"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = "grey90", linewidth = 0.2)
  ) +
  xlim(0, max(long_counts$Reads) * 1.13)
```
## <sup>CpG-site Coverage</sup>

* This metric evaluates how completely each sample covers the predefined CpG Panel. 
* Completeness is calculated as the proportion (%) of targeted CpG sites from the panel
* that were successfully detected, defined as having at least one mapped read.
* Bars: Represent the count of successfully covered CpG sites per sample.
* Percentage labels: Directly reflect completeness relative to the total panel size.

```{r}
#| echo: false
#| label: total-read-count panel  
#| error: true
#| fig-cap: "**Figure 13**: CpG Coverage Completeness per Sample"

total_panel_sites   <- nrow(panel)   
covered_panel_sites <- colSums(input.coverage > 0, na.rm = TRUE)

coverage_df <- tibble(
  Sample              = names(covered_panel_sites),
  Covered_Panel_Sites = covered_panel_sites
) |>
  mutate(
    Covered_Pct = 100 * Covered_Panel_Sites / total_panel_sites
  ) |>
  arrange(desc(Covered_Panel_Sites))                    

ggplot(coverage_df,
       aes(x = reorder(Sample, Covered_Panel_Sites),
           y = Covered_Panel_Sites)) +
  geom_col(fill = "#4E79A7", width = 0.75,
           colour = "grey20", linewidth = 0.15) +
  geom_text(aes(label = sprintf("%.1f%%", Covered_Pct)),
            hjust = -0.15, size = 3) +
  coord_flip(clip = "off") +
  scale_y_continuous(
    limits = c(0, total_panel_sites * 1.05),
    breaks = scales::pretty_breaks(5)
  ) +
  labs(
    title = "Panel CpG Sites Covered per Sample",
    subtitle = glue::glue("{total_panel_sites} total panel CpG sites"),
    x = "Sample",
    y = "CpG Sites"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.major.y = element_blank()
  )
```
## <sup>Coverage Uniformity</sup>

* `What the plot tests`. Each dot represents one sample;
* we contrast the mean and median depth across all targeted bases.
* `Uniform libraries hug the diagonal`. Points near the red dashed line (Mean ≈ Median)
* indicate homogeneous coverage, which is the goal.
* `Points below the line reveal skew`. Here, a handful of “hot-spot” regions with very high depth
* pull the mean upward, while most targets cluster around the lower median.
* `Why the median can be zero`. Picard’s
* [CollectHsMetrics](https://gatk.broadinstitute.org/hc/en-us/articles/360036856051-CollectHsMetrics-Picard)
* sets MEDIAN_TARGET_COVERAGE = 0 whenever ≥ 50 % of targets receive no passing reads (depth = 0).
* <em>In that case the dot collapses onto the X-axis,</em>
* <em>although the mean may remain > 0 thanks to the well-covered hot-spots</em>.


```{r}
#| echo: false
#| label: Assessment of Sequencing Coverage Uniformity
#| error: true
#| fig-cap: "**Figure 14**:This scatter plot compares the mean and median target coverage for each sample to evaluate the homogeneity of the sequencing data. The dashed diagonal line represents a perfect 1:1 ratio, indicating a completely uniform coverage distribution."
ggplot(hs,
       aes(`Mean target coverage`,
           `Median target coverage`,
           colour = Sample)) +
  geom_point(size = 2.2, alpha = .85,
             position = position_jitter(width = .02, height = 0)) +   
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "red") +
  scale_colour_discrete(labels = \(x) sub(".*_(S\\d+)", "\\1", x)) +
  labs(title  = "Median vs Mean Target Coverage",
       x      = "Mean target coverage",
       y      = "Median target coverage",
       colour = "Sample") +
  coord_equal() +                       
  theme_minimal(base_size = 13) +
  theme(legend.position = "right")
```
